A game-theory modeling approach to utility and strength of interactions dynamics in biomedical research social networks

### 2016-11-27 manuscript_3.tex
caught some "attachments"

Collaboration has become a cornerstone in biomedical research today.
In contrast to physics which has a long history and experience in
collaborative projects, biology is only recently becoming an evermore
collaborative discipline.
In this article we explore the effect of a collaboraion network on the distribution
of players having access to certain amount of resources from other players in the
network and the distribution of the strength of connections among
them. Particularly, we implemented two games played simultaneously:
one for maximazing individual utility based on the iterated Prisoner's
Dilemma; the other, a coordination game for maximazing the connection strength between
players. We are interested in how they affect each other in 
the context of a network of scientific collaboration under the idea that while
researchers are interested in maximazing their utilities, they also know that
it is important to invest in building collaborative relationships.
We tested our simulation on a biomedical research community network of M\'exico. and
compared the results with an Erd\"{o}s-Reny\'i, a Watts-Strogatz
small-world and Barab\'asi-Albert topologies. Different topologies
display different utility and attachment
distributions. Moreover, the distribution of utility and attachment in the
researchers network is similar to that of 
Barab\'asi-Albert and Watts-Strogatz topologies, respectively. We
believe that utility distribution in the researchers network
suggests that there are socio-cultural mechanisms governing the
network that produce an asymmetric distribution of resources. The
high distribution of strong connections might reflect some sort of
subordination among researchers by which they are morally obliged to
cooperate by the same socio-cultural mechanisms. The range around
the threshold that regulates the decision to cooperate or defect
according to the agent's historical balance between utility and
strength of collaborative relationships and carrying capacity of the
system is small, suggesting that there is a region in which a phase
transition takes place from a population of cooperators to a
population of defectors. Simulations like this may help to develop
science policies to promote fair distribution of resources.

>>>>

Collaboration has become a cornerstone in biomedical research today.
In contrast to physics which has a long history and experience in
collaborative projects, biology is only recently becoming an evermore
collaborative discipline.  In this article we explore the effect of
collaboraion on the distribution of players' access to resources from
other players in the network and the distribution of the strength of
interactions among them. Particularly, we implemented two games played
simultaneously: one for maximazing individual utility based on the
iterated Prisoner's Dilemma; the other, a coordination game for
maximazing the connection strength between players. We are interested
in how they affect each other in the context of a network of
scientific collaboration under the idea that while researchers are
interested in maximazing their individual utilities, they also know
that it is important to invest in building collaborative
relationships.  We tested our simulation on a biomedical research
community network of M\'exico and compared the results with an
Erd\"{o}s-Reny\'i, a Watts-Strogatz small-world and Barab\'asi-Albert
topologies. Different topologies display different utility and
interaction strength distributions. Moreover, the distribution of
utility and interaction strength in the researchers network is similar
to that of Barab\'asi-Albert and Watts-Strogatz topologies,
respectively. We believe that utility distribution in the researchers
network suggests that there are socio-cultural mechanisms governing
the network that produce an asymmetric distribution of resources. The
high distribution of strong interactions might reflect some sort of
subordination among researchers by which they are morally obliged to
cooperate by the same socio-cultural mechanisms. The range around the
threshold that regulates the decision to cooperate or defect according
to the agent's historical balance between utility and strength of
collaborative relationships and carrying capacity of the system is
small, suggesting that there is a region in which a phase transition
takes place from a population of cooperators to a population of
defectors. Simulations like this may help to develop science policies
to promote fair distribution of resources.
